---
title: "Claytor Solo Exercise"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

## Tidyverse Solo Exercise

In this project, you'll practice working with data using the tidyverse
libraries. You'll be working with data on each of 145 school districts
and the State of Tennessee. This data contains, for the 2014-2015 school
year: \* Proficiency rates on state tests \* Student demographics \*
Chronic absenteeism \* Discipline (suspension, expulsion) rates \* High
school graduation, dropout rates \* Average ACT composite scores \* A
region in Tennessee

Create an RMarkdown file to answer the following questions.

1.  Read in `districts.csv` into a tibble named `districts`. How many
    rows and columns does it contain?
    -   It contains 27 rows and 146 Columns

```{r load districts}
 districts
```


#2.  Notice that the first row corresponds to the whole State of
    Tennessee. Remove this row and save the result back to `districts`.
  
```{r remove first row.  DON'T PUSH AGAIN!!!}
  districts <- districts[-1,]
  districts
```
3.  How many districts have a proficiency rate of at least 80% for both
    alg_1 and eng_1? 

There were 13 districts where the proficiency for both Algebra 1 and English 1 were at-least 80%

```{r Freshman EOC 80% For Algebra 1 and English 1}
  freshEOC80 <- districts %>% 
    filter(alg_1 >=80 & eng_1>=80)
  freshEOC80
```

4.  How many districts have a proficiency rate less than 50% for either
    alg_1 or eng_1?
    
There were 8 counties where proficiency rates in either algebra 1 OR english 1 were below 50%

```{r Districts with freshaman ELA or Math Proficiency below 50%, paged.print=FALSE}
freshEOC50 <- districts %>% 
  filter(alg_1 <=50 | eng_1<=50)
freshEOC50
```


5.  Which district has the lowest graduation rate?
Achievement School District had the lowest graduation rate
```{r Lowest Graduation Rate}
LowestGraduation <- districts %>% 
  filter(grad == min(grad, na.rm=TRUE))
LowestGraduation
```

6.  Within the Mid Cumberland region, which district has the highest ACT
    composite?

Williamson County had the Highest ACT Composite from the Mid Cumberland Region.

```{r Highest ACT Composite in Midcumberland Region}
MidCumACT <- districts %>% 
  filter(region == "Mid Cumberland") %>% 
  filter(act_composite == max(act_composite, na.rm=TRUE))
MidCumACT
```

7.  Create a histogram showing the distribution of graduation rates.
    What can you say about this distribution?
```{r}
districts %>% 
  ggplot(aes(x=grad)) +
  geom_histogram()
```

8.  Create a scatter plot to compare alg_1 proficiency rates to alg_2
    rates. What do you notice? Facet this plot by region. Does anything
    stand out when you facet the plots?
```{r}
districts %>% 
  ggplot(aes(x= alg_1, y=alg_2, color = region)) +
  geom_point()
  facit_wrap(~region)
```
    

9.  Create a bar chart showing the total enrollment by region. Which
    region has the highest total enrollment? Which has the smallest?
```{r}
districts %>% 
  ggplot(aes(x= region, y=enrollment)) +
  geom_bar()
```
10. When creating this bar chart you may have noticed that some
    districts have missing enrollment values. For how many districts is
    this the case?

11. What is the mean graduation rate across all districts? What might be
    wrong with using just the regular mean to assess average graduation
    rates?

12. Redo the previous question but use a weighted average
    (`weighted.mean`) graduation across all districts, weighing by
    enrollment. How much does this change your answer? Can you explain
    using the data the reason for the big change from using the mean?

13. Create a boxplot showing graduation rates per region. Does anything
    stand out?

14. Find the weighted average of graduation rates by region using
    enrollment as weights. Compare the results you get for the weighted
    average to what you see from the boxplots.

15. For many districts, values for `alg_2` are lower than for `alg_1`.
    Create a histogram showing the distribution of differences
    (`alg_1` - `alg_2`). Which school had the largest drop from `alg_1`
    to `alg_2`? For what percentage of schools is it true that `alg_2`
    is larger than `alg_1`? Is there a similar drop off for `eng_2` and
    `eng_3`?

16. You may have noticed that a lot of rows are missing values. Which
    district has the largest number of missing values? What do you
    notice about schools that have a lot of missing values?

17. Find the correlation between graduation rate and all other
    variables. Create a horizontal bar chart showing these correlations.
    Make sure that your plot is ordered by correlation values. What do
    you notice from these correlations?

18. Create a scatterplot for `grad` vs. `suspended`. Does what you see
    make sense given your answer from the previous part?

19. Create a linear regression model using `lm` with target variable
    `grad` and predictor variable `suspended`. What R\^2 value does this
    model have? What is the interpretation of this number?

20. Add the regression line to your scatterplot using `geom_smooth` with
    `method='lm'`. How do you feel about the regression line after
    seeing it plotted on the scatterplot?

**Continued Exploration and Practice**

21. Read in the school-level testing data for 2014, available
    [here](https://www.tn.gov/content/dam/tn/education/data/data_2014_school_base.xlsx).
    You might find the readxl library useful for this task. If you use
    this library, be sure to look at the `na` argument for the
    `read_excel` function.

22. How many schools have at least 20 percent of students below bsc for
    Algebra I? Which districts do these schools belong to?

23. How many schools have at least 20 percent of students below bsc for
    *both* Algebra I and English I?

24. Which grade has the highest pct_adv for Algebra I? Plot the average
    pct_adv per grade level as a bar chart. Make sure that the bars are
    ordered by grade level.

25. Find the correlation between pct_adv for Algebra I and pct_adv for
    Algebra II by school. Create a scatterplot showing Algebra II scores
    vs. Algebra I scores by school.

26. Find all schools in Rutherford County that have "High School" in
    their name. For these schools, create a chart (your choice) showing
    the differences in pct_below_bsc, pct_bsc, pct_prof, and pct_adv for
    Algebra I when looking across all subgroups and grades.

27. I claim that smaller schools do a better job preparing students for
    Algebra I standardized tests. Find the average number of valid tests
    (a proxy for the school size) for schools where the pct_prof_adv for
    Algebra I is greater than 95. Compare this to the average number of
    valid tests for all schools. In light of this result, how does my
    claim look?

28. I also claim that smaller schools do a worse job preparing students
    for Algebra I standardized tests. Find the average number of valid
    tests (a proxy for the school size) for schools where the
    pct_prof_adv for Algebra I is less than 25. Compare this to the
    average number of valid tests for all schools. In light of this
    result, how does my claim look now?

29. Create a scatterplot showing pct_prov_adv vs. valid_tests. Can you
    use this to explain the result for the previous two questions?

If you finish all of the above questions, continue to explore the two
datasets and see what else interesting you can find.

Also, check out the plotly library for R. The `ggplotly` function makes
it very easy to convert ggplot plots into interactive plotly plots.
